The Evolution of Phenotypically Invariant Genetic Networks

### 2017-07-02 handnotes.tex
rescaled F2 hybrid figure
edited talk
made a poster

\begin{abstract}
 % \textbf{
I will outline an analytical theory to study the evolution of biological systems such as gene regulatory networks, borrowing insight and tools from control engineering, systems identification, and dynamical systems theory. I will describe a null model of regulatory network evolution by analytically describing the set of all linear gene networks (of any size) that produce identical phenotypes -- and the evolutionary paths connecting them. In the idealized case of a perfectly adapted population, constant selection, and a static environment, we observe neutral evolution as a random walk over the phenotypically- invariant network-space. Under neutral conditions, this model can provide descriptions of expected network size and connectivity under mutation-selection equilibrium, estimate the rate of regulatory rewiring, and the rates at which Dobzhansky-Muller incompatibilities arise in reproductively isolated populations. This analysis provides insight into the mechanisms and parameters important for understanding developmental systems drift, network rewiring, evolvability, epistasis, and speciation, as well as the tenuous connection between network architecture and function. 
%}
\end{abstract}

@@

      The fitness of an organism can be computed by comparing its impulse response with the optimal response,
      \plr{this is either zero or infinite.  apply a weighting function (see appendix)?}
      \begin{align}
        \mathcal{F}\left( \widehat{\phi}(t) \right) = \exp \left\{- \int_{0}^{\infty} \left\lVert \phi(t) - \widehat{\phi}(t) \right\rVert dt \right\}  .
      \end{align}
      Therefore a hybrid's fitness can be computed by comparing its impulse response with that of its parents.

@@

  %    \begin{figure}[H]
  %      \centering
  %      \includegraphics[width=0.5\textwidth]{A2_A2-1_F2s}
  %      \caption{F2s from $A(2)$ and $A(2.1)$.}
  %    \end{figure}
  %      \begin{figure}[H]
  %        \begin{tabular}{cc}
  %        \includegraphics[width=0.5\textwidth]{F2s-small} \\
  %        \includegraphics[width=0.5\textwidth]{F2s} \\
  %        \includegraphics[width=0.5\textwidth]{F2s-large}
  %        \end{tabular}
  %        \caption{F2s from $A(2)$ and $A(2.1)$.}
  %      \end{figure}
        \begin{figure}[H]
          \centering
          \begin{tabular}{cc}
            \includegraphics[width=0.5\textwidth, height=0.25\paperheight]{F1_comparison} & 
            \includegraphics[width=0.5\textwidth, height=0.25\paperheight]{F2_comparison}
          \end{tabular}
          \caption{F1 (left) and F2 (right) hybrids crossing $A(2)$ with $A(2.01)$ (top), $A(2.1)$ (middle), and $A(2.5)$ (bottom).
          Note the difference in scale on the $y$-axis. F2 hybrids display more phenotypic divergence than F1s, on average. Further, some F2s completely fail to oscillate, as seen in an $A(2.5)$ F2 (light blue).}
        \end{figure}
    \end{example}
      \begin{example}[Not all Networks can Host Incompatibilities]
        convex sets cant have DMIs

>>>>

\begin{abstract}
 % \textbf{
I will outline an analytical theory to study the evolution of biological systems such as gene regulatory networks, borrowing insight and tools from control engineering, systems identification, and dynamical systems theory. I will describe a null model of regulatory network evolution by analytically describing the set of all linear gene networks (of any size) that produce identical phenotypes – and the evolutionary paths connecting them. In the idealized case of a perfectly adapted population, constant selection, and a static environment, we observe neutral evolution as a random walk over the phenotypically invariant network-space. Under neutral conditions, this model can describe the rates at which Dobzhansky-Muller incompatibilities arise in reproductively isolated populations. This analysis provides insight into the mechanisms and parameters important for understanding developmental systems drift, network rewiring, and speciation.
%}
\end{abstract}

@@

      The fitness of an organism can be computed by comparing its impulse response with the optimal response,
      \plr{this is either zero or infinite.  apply a weighting function (see appendix)?}
      \begin{align}
        \mathcal{F}\left( \widehat{\phi}(t) \right) = \exp \left\{- \frac{1}{\sigma} \int_{0}^{\infty} \left\lVert \phi(t) - \widehat{\phi}(t) \right\rVert^{2} dt  \right\}  .
      \end{align}
      Therefore a hybrid's fitness can be computed by comparing its impulse response with that of its parents.

@@

  %    \begin{figure}[H]
  %      \centering
  %      \includegraphics[width=0.5\textwidth]{A2_A2-1_F2s}
  %      \caption{F2s from $A(2)$ and $A(2.1)$.}
  %    \end{figure}
  %      \begin{figure}[H]
  %        \begin{tabular}{cc}
  %        \includegraphics[width=0.5\textwidth]{F2s-small} \\
  %        \includegraphics[width=0.5\textwidth]{F2s} \\
  %        \includegraphics[width=0.5\textwidth]{F2s-large}
  %        \end{tabular}
  %        \caption{F2s from $A(2)$ and $A(2.1)$.}
  %      \end{figure}
        \begin{figure}[H]
          \centering
          \begin{tabular}{cc}
            \includegraphics[width=0.5\textwidth, height=0.25\paperheight]{F1_comparison} & 
            \includegraphics[width=0.5\textwidth, height=0.25\paperheight]{F2s_comparison2}
          \end{tabular}
          \caption{F1 (left) and F2 (right) hybrids crossing $A(2)$ with $A(2.01)$ (top), $A(2.1)$ (middle), and $A(2.5)$ (bottom).
          Note the difference in scale on the $y$-axis. F2 hybrids display more phenotypic divergence than F1s, on average. Further, some F2s completely fail to oscillate, as seen in an $A(2.5)$ F2 (light blue).}
        \end{figure}
    \end{example}
      \begin{example}[Not all Networks can Host Incompatibilities]
        convex sets cant have DMIs
